---
title: "KP_logRank"
author: "Qihun Yang"
date: "`r Sys.Date()`"
output: pdf_document
---

## Data Preparation
```{r pre}
#install.packages("survival")
library(survival)
setwd("D:\\Code_Factory\\surv_anly\\1")
ACTG175 <- read.table("ACTG175(speff2trial).txt", header = TRUE, sep = ',')
# used to indicate that all subsequent variable names apply to the ACTG175 dataset
attach(ACTG175)
```

## Q1: 
1. Plot.

(a) Draw a survival curve plot using the Kaplan-Meier method for the four treatment groups (i.e., \verb"arms"; $0=$ zidovudine, $1=$ zidovudine and didanosine, $2=$ zidovudine and zalcitabine, $3=$ didanosine). Compare the survival curves. What do you conclude?

Use functions \verb"survfit" and \verb"plot".
```{r Q1.1}
km <- Surv(days, cens==1)
# summary(survfit(km~arms))
plot(survfit(km~arms),ylim = c(0.6,1), col=c("red","orange","blue","green"),
     ylab = "survival probabilities", xlab="survival time in days")

legend("topright",           #The legend is in the upper right corner
       legend=c("zidovudine","zidovudine and didanosine","zidovudine and zalcitabine",
                "didanosine"),        #content of legend
       col=c("red","orange","blue","green"),         #color of legend
       lty=1,lwd=2,                                  #size of legend
       cex = 0.7, text.width = 380)
```
The plot indicates that subjects taken zidovudine have the significantly lowest rate of survival than other subjects. Subjects taken didanosine alone have a lower rate of survival than those combined didanosine with zidovudine. Two subjects seperately taken zidovudine and didanosine, zidovudine and zalcitabine have no significant difference.

(b) From the Kaplan-Meier survival curve plot, can you obtain the median survival time for each of the four treatment groups, respectively? Why?

No, we can't. Since the survival probabilities do not reach $0.5$ for all 4 subjects, Kaplan-Meier method is unable to do out-of-sample prediction.

## Q2:
2. Log-rank test.
(a) Suppose we are interested in testing whether the four treatment groups have the same effect on the survival outcome. To accomplish this goal, we conduct a log rank test. What is the $p$-value of the log rank test? What do you conclude?

Use functions \verb"survdiff" and \verb"plot".
```{r Q2.1.1}
#non-weighted log-rank test
survdiff(km~arms)
```

The log rank statistic is highly significant with a p-value of
$1\times10^{-10}$. Further, we can conduct log-rank test on $\verb"arms"=1,2,3$ whose similarity is obvious on the curves.
```{r Q2.1.2}
# use subset = group %in% c("group1", "group2") to do test only for a subset
survdiff(km~arms, subset = arms %in% c(1,2,3))
```

The log rank statistic is not significant with a p-value of
$0.4$. Thus, I deem that these three survival probabilities are identical.


(b) Conduct a weighted log rank test that emphasizes early survival difference. What is the p-value of the weighted log rank test? What do you conclude?

```{r Q2.2.1}
#weighted log-rank test
survdiff(km~arms, rho = 1)
```
The results of the test in which rho=1 yield a chi-square
value of $53$ with a p-value of $2\times10^{-11}$, which shows a
more significant effect of \verb"arms" on survival the non-weighted version.

(c) Conduct a stratified log rank test, where we stratify the variables \verb"hemo", \verb"homo", and \verb"age" simultaneously. In particular, we divide \verb"age" into three categories " $\leq 20$ years old", "20 - 50 years old" and ">50 years old". What is the p-value of this stratified log rank test? What do you conclude?
```{r Q2.3.1}
#stratified log-rank test
survdiff(km~arms + strata(hemo) + strata(homo) +
           strata(cut(age, breaks = c(0,20, 50,max(age)))))
```
I use the function \verb"cut" to convert numeric variable \verb"age" into factor.

The p-value of this stratified test is $2\times10^{-11}$, which is one fifth of the test without stratifying \verb"hemo", \verb"homo" and \verb"age" but still obtaining the same conclusion.
